A UK Case Study on\\Cybersecurity Education and Accreditation

### 2019-07-26 FIE2019-final.tex
Changes to Methodology section

\section{Research Questions}

There are various levels of specialism at which cybersecurity education can be addressed:
\begin{enumerate}[label=(\roman*)]
\item The general public --- this is important, but there are many initiatives in this area, which are, rightly, largely separated from computing education;
% ({\emph{interesting point from a public engagement/science communication perspective -- aren't ``we'' trying to create a digitally competent and capable citizenry? Minor point but maybe worth mentioning e.g. being safe online, accessing public services, etc}})
%\item[JHD]This is important, agreed, but I think this is being a user: think car driving rather than car design. BUT the point I heard from the WC Engineers ``how can security be an option --- safety certainly isn't'' is also relevant.
\item The generalist computer science graduate;
\item The generalist computer science masters graduate;
\item The specialist computer science graduate;
\item The specialist computer science masters graduate.
\end{enumerate}

The focus of this paper is on (ii)--(iii): the general computer science graduate. We thus focus on three research questions:

@@

A UK-focused case study-based approach is adopted in this project. As is common in case study-based research, many alternative case studies could have been chosen. The cases are illustrative rather than comprehensive in terms of the available case studies or challenges. The cases were evaluated to articulate the progress made and highlight opportunities for future developments in education and practice.

The first set of cases that are considered are the current situation in terms of recommendations from a sample of relevant PSRBs, together with the published evidence regarding compliance with these recommendations. This is followed by the evaluation of the policy context. Together these provide a context to the ongoing enhancement initiatives in the area of cybersecurity education.

In order to address {\emph{RQ1}} and {\emph{RQ2}} a number of case studies pertaining to challenges of delivering cybersecurity to undergraduate computer science students are evaluated. Namely, an industry problem with evident cybersecurity implications; the current state of educational resources with respect to cybersecurity; and the challenges evident in the UK related to the recruitment and retention of suitably qualified academic staff in the cyber security area.

Finally, in order to address {\emph{RQ3}} cases studies are evaluated related to the challenges and successes of PSRB accreditation of cybersecurity in undergraduate computer science in one jurisdiction (the UK). As discussed in the current situation section, in this jurisdiction mandating cybersecurity within PSRB accreditation is at a reasonably mature stage.

@@

The ACM/IEEE-CS Joint Task Force on Computing Curricula~\cite[p.~97]{ACM2013a} takes a distinct view on the Knowledge Areas (KAs) related to RQ2:
\begin{quote}
{\emph{The Information Assurance and Security KA is unique among the set of KAs presented here
given the manner in which the topics are pervasive throughout other KAs}}
\end{quote}
It proposes 9 ``core'' hours and 63.5 distributed across the other KAs.
%\pagebreak - to fix URL issues
\par
Nevertheless, the situation on the ground in the USA is different~\cite{Ackerman2019a}:
\begin{quote}
{\emph{Universities suffer shortcomings, as well. Roughly 85 of them offer undergraduate and/or graduate degrees in cybersecurity. There is a big catch, however. Far more diversified computer science programs, which attract substantially more students, don't mandate even one cybersecurity course.}}
\end{quote}

The UK situation is distinctly different: 61\% of UK courses offer mandatory cybersecurity content, and this research was based on web scraping~\cite[Table 1]{Ruiz2019a}. As such it represents a lower bound since not all coverage will necessarily be clearly articulated in publicly available documentation online.

It is at least plausible to attribute this difference to differences in the accreditation regimes, as the external circumstances, governmental pressures, and professional body/learned society curricula are all similar.

\begin{description}
\item[UK]BCS, The Chartered Institute for IT (BCS) has had a requirement to include information security in the curriculum since 2010, and has expected coverage of an agreed cybersecurity syllabus since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (because of the five-year cycle). More precisely, accredited degrees have been expected to demonstrate coverage of ``{\emph{2.1.9 Knowledge and understanding of information security issues in relation to the design, development and the use of information systems}}'' \cite[p.~30]{BCS2018a} since 2010 with an enhanced cybersecurity related definition of what this entails since 2015 \cite[p.~17--18]{BCS2018a}.
%BCS has had cybersecurity in the curriculum since 2013, and required cybersecurity in accreditation since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (because of the five-year cycle).
\item[US]ACM has equally had cybersecurity (``Information Assurance and Security'' --- IAS) in the curriculum since 2013 \cite{ACM2013a}, but it is not the accrediting body. ABET is, and is requiring IAS with effect from the 2019-20 cycle (self-study reports due 1 July 2019): more precisely \cite[Table 3]{Oudshoornetal2018a} ``{\emph{The computing topics must include: \dots{} Principles and practices for secure computing\dots}}''. This should mean that  all accredited universities should be compliant by 2025 (because of the six-year cycle).
%\cite[chapter 7]{Davenportminor2019a}.  % need a better reference
\end{description}

@@

%What should be in the general curriculum.  Pervasive (as in Maths or LSEPI) rather than specialist option.  
%{\bf JHD to research ACM general curriculum;}
The ACM Computer Science Curricula Recommendations~\cite{ACM2013a} states three Tier-1 and six Tier-2 hours for ``Information Assurance and Security'', but this is the ``IAS-only'' topics, and ACM expects 32 Tier-1 and 31.5 Tier-2 Hours for IAS topics embedded in other Knowledge Areas.

>>>>

\section{Methodology}

\subsection{Research Questions}

There are various levels of specialism at which cybersecurity education and skills can be addressed; for example:
\begin{enumerate}[label=(\roman*)]
% ({\emph{interesting point from a public engagement/science communication perspective -- aren't ``we'' trying to create a digitally competent and capable citizenry? Minor point but maybe worth mentioning e.g. being safe online, accessing public services, etc}})
%\item[JHD]This is important, agreed, but I think this is being a user: think car driving rather than car design. BUT the point I heard from the WC Engineers ``how can security be an option --- safety certainly isn't'' is also relevant.
\item The generalist computer science graduate;
\item The generalist computer science masters graduate;
\item The specialist computer science graduate;
\item The specialist computer science masters graduate;
\item The reskilling/upskilling/professional development of the IT industry and the wider workforce;
\item The general public --- this is important, but there are many initiatives in this area, which are, rightly, largely separated from computing education.
\end{enumerate}

The focus of this paper is on (i)--(ii): the general computer science graduate. We thus focus on three research questions:

@@

A UK-focused case study-based approach has been adopted for this project. As is common in case study-based research, many alternative case studies could have been chosen; furthermore, the cases are illustrative rather than comprehensive in terms of the available case studies or challenges. The cases were evaluated to articulate the progress made and highlight opportunities for future developments in education and practice.

The first set of cases that are considered are the current situation in terms of recommendations from a sample of relevant PSRBs, together with the published evidence regarding compliance with these recommendations; this is contextualised by the wider policy context. Together these provide a context to the ongoing enhancement initiatives in the area of cybersecurity education.

In order to address {\emph{RQ1}} and {\emph{RQ2}} a number of specific case studies pertaining to challenges of delivering cybersecurity to undergraduate computer science students are evaluated. Namely: an industry problem with evident cybersecurity implications; the current state of educational resources with respect to cybersecurity; and the challenges evident in the UK related to the recruitment and retention of suitably qualified academic staff in the cyber security area.

Finally, in order to address {\emph{RQ3}}, case studies are evaluated related to the challenges and successes of PSRB accreditation of cybersecurity in undergraduate computer science in one jurisdiction (the UK). As discussed, mandating cybersecurity within PSRB accreditation in this jurisdiction is at a reasonably mature stage.

@@

The ACM/IEEE-CS Joint Task Force on Computing Curricula~\cite[p.~97]{ACM2013a} takes a distinct view on the Knowledge Areas (KAs) related to {\emph{RQ2}}:

\begin{quote}
``{\emph{The Information Assurance and Security KA is unique among the set of KAs presented here
given the manner in which the topics are pervasive throughout other KAs.}}''
\end{quote}

It proposes nine ``core'' hours and 63.5 distributed across the other KAs. Nevertheless, the situation on the ground in the USA is different~\cite{Ackerman2019a}:

\begin{quote}
``{\emph{Universities suffer shortcomings, as well. Roughly 85 of them offer undergraduate and/or graduate degrees in cybersecurity. There is a big catch, however. Far more diversified computer science programs, which attract substantially more students, don't mandate even one cybersecurity course.}}''
\end{quote}

The UK situation is distinctly different: 61\% of UK courses offer mandatory cybersecurity content, and this research was based on web scraping~\cite[Table 1]{Ruiz2019a}. As such it represents a lower bound since not all coverage will necessarily be clearly articulated in publicly available documentation online. It is at least plausible to attribute this difference to differences in the accreditation regimes, as the external circumstances, governmental pressures, and professional body/learned society curricula are all similar.

\begin{description}
\item[UK] BCS, The Chartered Institute for IT (BCS) has had a requirement to include information security in the curriculum since 2010, and has expected coverage of an agreed cybersecurity syllabus since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (due to the five-year cycle). More precisely, accredited degrees have been expected to demonstrate coverage of ``{\emph{2.1.9 Knowledge and understanding of information security issues in relation to the design, development and the use of information systems}}'' \cite[p.~30]{BCS2018a} since 2010 with an enhanced cybersecurity related definition of what this entails since 2015 \cite[p.~17--18]{BCS2018a}.
\item[USA] The Association of Computing Machinery (ACM) has equally had cybersecurity (IAS: ``Information Assurance and Security'') in the curriculum since 2013~\cite{ACM2013a}, but it is not the accrediting body. The Accreditation Board for Engineering and Technology (ABET) is, and is requiring IAS with effect from the 2019-20 cycle (self-study reports due 1 July 2019): more precisely \cite[Table 3]{Oudshoornetal2018a} ``{\emph{The computing topics must include: \dots{} Principles and practices for secure computing\dots}}''. This should mean that  all accredited universities should be compliant by 2025 (due to their six-year cycle).
\end{description}

@@

%What should be in the general curriculum.  Pervasive (as in Maths or LSEPI) rather than specialist option.  
The ACM Computer Science Curricula Recommendations~\cite{ACM2013a} states three Tier-1 and six Tier-2 hours for ``Information Assurance and Security'', but this is the ``IAS-only'' topics, and ACM expects 32 Tier-1 and 31.5 Tier-2 Hours for IAS topics embedded in other Knowledge Areas.
